Practical Optional Types for Clojure

### 2015-02-19 formal-model.tex
use minted for syntax highlighting
do example

>>>>

\subsection{\doliteral{}}

Assertions in Clojure often expand to simple exceptional control flow reasoning.

\begin{minted}{clojure}
(fn [x :- (U nil Number)] {:pre (number? x)} (inc x))
\end{minted}

roughly expands to

\begin{minted}{clojure}
(fn [x :- (U nil Number)]
  (do (if x (throw (new Exception)) nil)
      (inc x)))
\end{minted}

With an elegant extension in T-Do (figure~\ref{figure:typingrules}) we can follow this reasoning
to infer \mintinline{clojure}{(inc x)} is safe. The main insight is we can assume the
disjunction of the propositions of the first \doliteral{} argument while checking the second.

To sketch how this example is checked, first T-Abs type checks the body under the 
environment \propenv{} = {\isprop{\Union{\nil}{\Number{}}}{\x{}}}.
T-Do then extracts the propositions of its first argument:
T-If tells us \thenprop{\prop{1}} = {\orprop{\botprop{}}{\botprop{}}} and 
\elseprop{\prop{1}} = {\orprop{\botprop{}}{\notprop{\falsy{}}{\x{}}}}
using T-Error, T-Nil and T-Subsume.
Now {\orprop {\thenprop {\prop{1}}} {\elseprop {\prop{1}}}} = {\notprop{\falsy}{\x{}}}
and we check \mintinline{clojure}{(inc x)} in the extended environment 
\ma{{\propenv{},{\notprop{\falsy}{\x{}}}}}
and by L-Update we know {\isprop{\Number}{\x{}}}
so T-Local says \mintinline{clojure}{x} is a \Number.
